---
title: "lesson4_problem_set"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Set up diamond data

```{r}
library(ggplot2)
data("diamonds")
summary(diamonds)
```

###Quiz 1

```{r}
dim(diamonds)

is.factor(diamonds$depth)
is.ordered((diamonds$carat))

?diamonds
```

###Quiz 2

```{r}
qplot(data = diamonds, x = diamonds$price, binwidth = 100,
      color = I('black'), fill = I("blue"))
```

###Quiz 3

```{r}
summary(diamonds$price)
```
The graph is long-tailed and skewed right so the center is towards the left. To be exact:

   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    326     950    2401    3933    5324   18823 
    
###Quiz 4

```{r}
dim(subset(diamonds, price < 500))
dim(subset(diamonds, price < 250))
dim(subset(diamonds, price >= 15000))
```

###Quiz 5

```{r}
qplot(data = diamonds, x = diamonds$price, binwidth = 100,
      color = I('black'), fill = I("blue")) + 
  scale_x_continuous(limits = c(0, 2500), breaks = seq(0, 2500, 500))
```

###Quiz 6

```{r}
cut1 = qplot(data = subset(diamonds, diamonds$cut == 'Fair'), x = price)
cut2 = qplot(data = subset(diamonds, diamonds$cut == 'Good'), x = price)
cut3 = qplot(data = subset(diamonds, diamonds$cut == 'Very Good'), x = price)
cut4 = qplot(data = subset(diamonds, diamonds$cut == 'Premium'), x = price)
cut5 = qplot(data = subset(diamonds, diamonds$cut == 'Ideal'), x = price)
grid.arrange(cut1, cut2, cut3, cut4, cut5, ncol = 2)
```

###Quiz 7

```{r}
by(diamonds$price, diamonds$cut, summary)
```

###Quiz 8

```{r}
qplot(x = price, data = diamonds) + facet_wrap(~cut)

#makes the scales free in the y-axis
qplot(x = price, data = diamonds) + facet_wrap(~cut, scales = "free_y")
```

###Quiz 9

```{r}
qplot(x = price/carat, data = diamonds) + 
  facet_wrap(~cut, scales = "free_y") +
  scale_x_log10()

```

###Quiz 10

```{r}
by(diamonds$price, diamonds$clarity, summary)
by(diamonds$price/diamonds$carat, diamonds$clarity, summary)

p1 = qplot(data = diamonds, x = clarity, y = price, geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 6000))

p2 = qplot(data = diamonds, x = clarity, y = price/carat, geom = 'boxplot') +
  coord_cartesian(ylim = c(2000, 6000))

grid.arrange(p1, p2)
ggsave("priceHistogram.png")
```

###Quiz 11

```{r}
by(diamonds$price, diamonds$color, summary)
```

###Quiz 12

```{r}
p1 = qplot(x = color, y = price, data = diamonds, geom = "boxplot") +
  coord_cartesian(ylim = c(0, 8000))

p2 = qplot(x = color, y = price/carat, data = diamonds, geom = "boxplot") +
  coord_cartesian(ylim = c(2000, 6000))

ggsave("price_by_color_hist.png")
by(diamonds$price/diamonds$carat, diamonds$color, summary)
```

###Quiz 13

```{r}
qplot(x = carat, y = ..count.., data = diamonds, geom = 'freqpoly', 
      binwidth = 0.1,
      ylab = "Number of Diamonds",
      xlab = "Carats") +
  scale_x_continuous(lim = c(0, 2.5), breaks = seq(0, 2.5, 0.1)) +
  scale_y_continuous(lim = c(0, 11000), breaks = seq(0, 11000, 1000))
```

###Quiz 14

```{r}
sugar <- read.csv("sugar_consumption.csv", header = T, row.names = 1, check.names = F)
transposed_sugar <- as.data.frame(t(sugar))
#column names
names(transposed_sugar)
#row names
row.names(transposed_sugar)
#remove all columns that are empty
transposed_sugar[sapply( X = transposed_sugar, FUN =  function(x) all(is.na(x)))] <- NULL
pt1 <- qplot(data = transposed_sugar, x = Philippines, binwidth = 5,
      color = I("black"), fill = I("blue")) + 
  scale_x_continuous(limits = c(0, 200))

pt2 <- qplot(data = transposed_sugar, x = Niger, binwidth = 5,
      color = I("black"), fill = I("red")) + 
  scale_x_continuous(limits = c(0, 200))

pt3 <- qplot(data = transposed_sugar, x = France, binwidth = 5,
      color = I("black"), fill = I("green")) + 
  scale_x_continuous(limits = c(0, 200))

#change united states name to remove space
names(transposed_sugar)[names(transposed_sugar)=="United States"] <- "USA"

pt4 <- qplot(data = transposed_sugar, x = USA, binwidth = 5,
      color = I("black"), fill = I("yellow")) + 
  scale_x_continuous(limits = c(0, 200))

grid.arrange(pt1, pt2, pt3, pt4)

summary(transposed_sugar$USA)
summary(transposed_sugar$France)
summary(transposed_sugar$Philippines)
summary(transposed_sugar$Niger)

```

#analysis
I chose four countries to compare sugar consumption. I picked countries from different continents. I was not surprised that Niger had a much lower sugar consumption than the other countries or the order of which countries had the least and most sugar consumption. The only thing that I found surprising was how much sugar in general was consumed. According to the American Health Association, males should consume less than 37.5 grams per day and women 25 grams. Only Niger is below that, and the other three countries are very much above.

```{r}
#sugar graph

#change sugar names 
names(sugar)[names(sugar)=="1961"] <- "x1961"
names(sugar)[names(sugar)=="1971"] <- "x1971"
names(sugar)[names(sugar)=="1981"] <- "x1981"
names(sugar)[names(sugar)=="1991"] <- "x1991"
names(sugar)[names(sugar)=="2001"] <- "x2001"

p1 = qplot(data = sugar, x = x1961, binwidth = 5,
      color = I("black"), fill = I("yellow")) +
  scale_x_continuous(limits = c(0, 200)) +
  scale_y_continuous(limits = c(0,20))

p2 = qplot(data = sugar, x = x1971, binwidth = 5,
      color = I("black"), fill = I("blue")) +
  scale_x_continuous(limits = c(0, 200)) +
  scale_y_continuous(limits = c(0,20))

p3 = qplot(data = sugar, x = x1981, binwidth = 5,
      color = I("black"), fill = I("red")) +
  scale_x_continuous(limits = c(0, 200)) +
  scale_y_continuous(limits = c(0,20))

p4 = qplot(data = sugar, x = x1991, binwidth = 5,
      color = I("black"), fill = I("orange")) +
  scale_x_continuous(limits = c(0, 200)) +
  scale_y_continuous(limits = c(0,20))

p5 = qplot(data = sugar, x = x2001, binwidth = 5,
      color = I("black"), fill = I("green")) +
  scale_x_continuous(limits = c(0, 200)) +
  scale_y_continuous(limits = c(0,20))

grid.arrange(p1, p2, p3, p4, p5)

summary(sugar$x1961)
summary(sugar$x1971)
summary(sugar$x1981)
summary(sugar$x1991)
summary(sugar$x2001)

```

#analysis
While sugar consumption has certainly increased incrementally each decade, the amounts have not been as high as I would have expected. I am interested in learning how sugar consumption has increased in developing countries, so as GDP increases, does sugar increase similarly? Maybe something to learn next lesson.

```{r}
#boxplot
sugar[sapply( X = sugar, FUN =  function(x) all(is.na(x)))] <- NULL
b1 = qplot(data = sugar, x = "1961", y = sugar$x1961,
      ylab = "Sugar Consumed in Grams",
      geom = 'boxplot') +
  scale_y_continuous(limits = c(0, 200))

b2 = qplot(data = sugar, x = "1971", y = sugar$x1971,
      ylab = "Sugar Consumed in Grams",
      geom = 'boxplot') +
  scale_y_continuous(limits = c(0, 200))

b3 = qplot(data = sugar, x = "1981", y = sugar$x1981,
      ylab = "Sugar Consumed in Grams",
      geom = 'boxplot') +
  scale_y_continuous(limits = c(0, 200))

b4 = qplot(data = sugar, x = "1991", y = sugar$x1991,
      ylab = "Sugar Consumed in Grams",
      geom = 'boxplot') +
  scale_y_continuous(limits = c(0, 200))

b5 = qplot(data = sugar, x = "2001", y = sugar$x2001,
      ylab = "Sugar Consumed in Grams",
      geom = 'boxplot') +
  scale_y_continuous(limits = c(0, 200))

grid.arrange(b1, b2, b3, b4, b5, ncol=5)
```

