---
title: "lesson8_problem_set"
output: html_document
---

###Quiz 1

```{r}
data(diamonds)

head(diamonds)

ggplot(aes(x = log(price)), data = diamonds) +
  facet_wrap(~color) +
  geom_histogram(aes(color = cut))
```

###Quiz 2

```{r}
ggplot(aes(x = table, y = price), data = diamonds) +
  geom_point(aes(color = cut)) +
  scale_x_continuous(limits = c(50, 80))
```

###Quiz 4

```{r}
diamonds.vol <- mutate(diamonds, volume = x * y * z)

ggplot(aes(x = volume, y = price), 
       data = diamonds.vol) +
  geom_point(aes(color = clarity)) +
  scale_x_continuous(limits = c(0, quantile(diamonds.vol$volume, .99))) +
  scale_color_brewer(type = 'div') +
  scale_y_log10()
```

###Quiz 5

```{r}
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)


pf <- mutate(pf, prop_initiated = friendships_initiated/friend_count)

tail(pf, 20)
```

###Quiz 6

```{r}
pf$year_joined <- floor(2014 - (pf$tenure/365))
pf$year_joined.bucket <- cut(pf$year_joined, 
                             breaks = c(2004, 2009, 2011, 2012, 2014))

table(pf$year_joined.bucket)

ggplot(aes(x = tenure, y = prop_initiated), 
       data = subset(pf, !is.na(year_joined.bucket))) + 
geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)

```

###Quiz 7

```{r}
ggplot(aes(x = 70*round(tenure/70), y = prop_initiated), 
       data = subset(pf, !is.na(year_joined.bucket))) + 
geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)

ggplot(aes(x = tenure, y = prop_initiated), 
       data = subset(pf, !is.na(year_joined.bucket))) + 
geom_smooth(aes(color = year_joined.bucket))
```

###Quiz 9

```{r}
table(pf$year_joined.bucket)
pf$year_joined.bucket[levels(pf$year_joined.bucket)[4]]

summary(subset(pf, pf$year_joined %in% (2013:2015)))
```

I believe that this group's proportion is higher for two reasons. First, when facebook first started, people would usually only initiate friendships with people they know well. More recently, people have begun to initiate friendships with casual acquantances as well.
Second, as we have seen, people tend to add a lot of friends when they first join facebook, so naturally the people in the most recent bin have just most recently had that initial burst of friend adding.

###Quiz 10

```{r}
ggplot(aes(x = cut, y = price/carat), data = diamonds) +
  geom_jitter(aes(color = color)) +
  facet_wrap(~clarity) +
  scale_color_brewer(type = 'div')
```

###Quiz 11

```{r}
install.packages('plyr')
library(plyr)

#unused
#PCs
pc <- read.csv("personal_computers_per_100.csv")
pc.gathered <- gather(pc, 'year', 'n', 2:20)
colnames(pc.gathered) <- c('country', 'year', 'pc')

#broadband
broadband <- read.csv("broadband_per_100.csv")
broadband.gathered <- gather(broadband, 'year', 'n', 2:15)
colnames(broadband.gathered) <- c('country', 'year', 'broadband')

#internet users
internet_users <- read.csv("internet_user_per_100.csv")
internet_users.gathered <- gather(internet_users, 'year', 'n', 2:23)
colnames(internet_users.gathered) <- c('country', 'year', 'internet_users')

#used
#bmi male
bmi_m <- read.csv("bmi_male.csv")
bmi_m.gathered <- gather(bmi_m, 'year', 'n', 2:30)
colnames(bmi_m.gathered) <- c('country', 'year', 'bmi_m')
head(bmi_m.gathered)

#bmi female
bmi_f <- read.csv("bmi_female.csv")
bmi_f.gathered <- gather(bmi_f, 'year', 'n', 2:30)
colnames(bmi_f.gathered) <- c('country', 'year', 'bmi_f')

#GDP growth
gdp_growth <- read.csv("gdp_growth.csv")
gdp_growth.gathered <- gather(gdp_growth, 'year', 'n', 2:46)
colnames(gdp_growth.gathered) <- c('country', 'year', 'gdp_growth')

#working hours
working_hours <- read.csv("hours_per_week.csv")
working_hours.gathered <- gather(working_hours, 'year', 'n', 2:30)
colnames(working_hours.gathered) <- c('country', 'year', 'working_hours')

#combine and clean dfs
dfs <- list(working_hours.gathered, gdp_growth.gathered, 
            bmi_m.gathered, bmi_f.gathered)
df <- join_all(dfs, by = c('country', 'year'), type='inner')
df <- subset(df, df$year != 'X')
df <- subset(df, df$country != '')
df$year <- gsub('X', '', df$year)

head(df)
tail(df)

df_select_countries <- filter(df, 
       df$country %in% c("United States", "Brazil", 
                                  "Germany", "Australia", "Japan"))

df_select_countries$country
```

```{r}
library(gridExtra)
#bmi
f <- ggplot(aes(x = year, y = bmi_f), 
       data = df_select_countries) + 
  geom_point(aes(color = country)) +
  scale_y_continuous(limits = c(21, 29))
m <- ggplot(aes(x = year, y = bmi_m), 
       data = df_select_countries) + 
  geom_point(aes(color = country)) +
  scale_y_continuous(limits = c(21, 29))

grid.arrange(f, m, ncol = 2)
```


```{r}
#bmis vs country
ggplot(aes(x = bmi_f, y = bmi_m), data = df_select_countries) + 
  geom_point(aes(color = country)) +
  scale_x_continuous(limits = c(21, 29)) +
  scale_y_continuous(limits = c(21, 29))
```

```{r}
#create year bins
df_select_countries <- subset(df_select_countries, !is.na(year))

df_select_countries <- mutate(df_select_countries, 
                              year.num = as.integer(year))
df_select_countries$year.num
head(df_select_countries)
df_select_countries$year.num.bucket <- cut(df_select_countries$year.num,
                                           breaks = c(1980, 1990, 2000, 2010))

table(df_select_countries$year.num.bucket)

table(df_select_countries$year.bucket)

df_select_countries$year.bucket

#bmi average vs gdp in year bins
ggplot(aes(x = gdp_growth, y = (bmi_f + bmi_m)/2), data = df_select_countries) +
  geom_point(aes(color = country)) +
  facet_wrap(~year.num.bucket)
```

```{r}
#working hours vs gdp
ggplot(aes(x = gdp_growth, y = working_hours), data = df_select_countries) +
  geom_point(aes(color = country))

ggplot(aes(x = gdp_growth, y = working_hours), data = df) +
  geom_point(aes(color = country))
```

