Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(dplyr)
library(ggplot2)

pf = read.csv('pseudo_facebook.tsv', sep = '\t')

age_gender_groups <- group_by(pf, age, gender)
pf.fc_by_age_gender <- summarise(age_gender_groups, 
                                 mean_friend_count = mean(friend_count), 
                                 median_friend_count = median(friend_count),
                                 n = n())

#another way to do it

pf.fc_by_age_gender2 <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>% 
  summarize(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>% 
  ungroup() %>% 
  arrange(age)
head(pf.fc_by_age_gender2)

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_histogram()


```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count), data = pf.fc_by_age_gender2) +
  geom_line(aes(color = gender))

#same graph different implementation
ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender2, 
                                  age ~ gender,
                                  value.var = 'median_friend_count')
head(pf.fc_by_age_gender2)
head(pf.fc_by_age_gender.wide)

#try to convert back to long format

pf.fc_by_age_gender.long <- melt(pf.fc_by_age_gender.wide, id='age')

head(pf.fc_by_age_gender.long)
tail(pf.fc_by_age_gender.long)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
pf.fc_by_age_gender.ratio <- pf.fc_by_age_gender.wide %>% 
  group_by(age) %>% 
  mutate(ratio = female/male) %>% 
  arrange(age)

head(pf.fc_by_age_gender.ratio)

ggplot(aes(x = age, y = ratio), data = pf.fc_by_age_gender.ratio) + 
  geom_line(color = 'blue') +
  geom_hline(yintercept = 1, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - (pf$tenure/365))

head(pf)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)

?cut

pf$year_joined.bucket <- cut(pf$year_joined, 
                             breaks = c(2004, 2009, 2011, 2012, 2014))
table(pf$year_joined.bucket)
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
table(pf$year_joined.bucket, useNA = 'ifany')

ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
subset(pf, tenure >= 1) %>% 
  mutate(friending_rate = friend_count/tenure) %>% 
  summarize(median = median(friending_rate), mean = mean(friending_rate),
            max = max(friending_rate))

#altenative solution
with(subset(pf, tenure >= 1), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated/tenure), 
       data = subset(subset(pf, !is.na(year_joined.bucket)), 
                     tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

#quiz answer
ggplot(aes(tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket))

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)

#change id fro int to factor
yo$id <- factor(yo$id)
str(yo)

ggplot(aes(x = price), data = yo) +
  geom_histogram(binwidth = 1, color = 'black', fill = 'blue') +
  scale_x_continuous(breaks = seq(0, 70, 5))
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
summary(yo)
unique(yo$price)
length(unique(yo$price))
table(yo$price)

yo <- mutate(yo, all.purchases = strawberry + blueberry + pina.colada +
         plain + mixed.berry)
head(yo)


```

***

### Prices over Time
Notes:

```{r Prices over Time}
head(yo)
#shape 21 is hollow circle
ggplot(aes(x = time, y = price), data = yo) +
  geom_jitter(alpha = 1/10, shape = 21, fill = 'orange')
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
#class seed
set.seed(4230)

#my seed
set.seed(1853)
sample.ids <- sample(levels(yo$id), 16)
sample.ids

ggplot(aes(x = time, y = price), 
       data = subset(yo, id %in% sample.ids)) +
  facet_wrap(~id) +
  geom_line() +
  geom_point(aes(size = all.purchases), pch = 1)

ggsave('household_sample.png')
```

In general, prices increase as time passes. A few of the households were very consistent in their purchases. These households purchased the majority of the yogurt. There were times that it was very likely they used coupons to get good prices on the yogurt. I found it interesting that on these days they usually didn't buy more yogurt than usual.

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix

```{r}
install.packages('GGally')
library(GGally)

pf_subset = pf[, c('age', 'dob_year', 'dob_month', 'gender', 'tenure')]
names(pf_subset)

ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])

```

Notes:

***

### Even More Variables
```{r}
nci <- read.table('nci.tsv')

colnames(nci) <- c(1:64)
```

Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

